---
title: "Cross Validation"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages.

```{r}
library(tidyverse)
library(p8105.datasets)
library(modelr)
```

Load the LIDAR

```{r}
data("lidar")
```

Look at the data

```{r}
lidar 

lidar_df = 
  lidar |> 
  mutate(id = row_number())

lidar_df |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point()
# non-constant variance
```

## Create dataframes

```{r}
train_df = 
  sample_frac(lidar_df, size = .8) |> 
  arrange(id)

test_df = anti_join(lidar_df, train_df, by = "id") # rows in lidar_df that aren't in the train_df
```


Look at these

```{r}
ggplot(train_df, aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")
# shows training and testing data
```

Fit a few models to `train_df`.

```{r}
# linear model that takes log ratio
linear_mod = lm(logratio ~ range, data = train_df)

# non-linear model fitting
smooth_mod = mgcv::gam(logratio ~ s(range), data = train_df)

# adding in parameters (k), don't do this because this model is overfitting
wiggly_mod = mgcv::gam(logratio ~ s(range, k = 50), sp = 10e-8, data = train_df)
```


Look at this!!

```{r}
# Fit linear model
train_df |> 
  add_predictions(linear_mod) |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
# Fit non-linear model
train_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

```{r}
# Fit wiggly model
train_df |> 
  add_predictions(wiggly_mod) |> 
  ggplot(aes(x = range, y = logratio)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

Try computing our RMSEs

```{r}
# testing models on testing dataset (want to minimize RMSE)
rmse(linear_mod, test_df) # not good
rmse(smooth_mod, test_df) # best model
rmse(wiggly_mod, test_df) # not good
```




## ITERATE!!

```{r}
cv_df = 
  # 80/20 split by default
  crossv_mc(lidar_df, n = 100) |> # doing split 100 times
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Did this work? Yes!

```{r}
# third training/testing split
cv_df |> pull(train) |> nth(3)
```

Let's fit models over and over.

```{r}
lidar_lm = function(df) {
  
  lm(logratio ~ range, data = df)
  
}

cv_df |>
  mutate(
    linear_fits = map(train, lidar_lm)
  )
```


```{r}
cv_df =
  cv_df |> 
  mutate(
    linear_fits = map(train, \(df) lm(logratio ~ range, data = df)),
    smooth_fits = map(train, \(df) mgcv::gam(logratio ~ s(range), data = df)),
    wiggly_fits = map(train, \(df) mgcv::gam(logratio ~ s(range, k = 50), sp = 10e-8, data = df))
  ) |> 
  mutate(
    # compute the RMSE for the different models
    rmse_linear = map2_dbl(linear_fits, test, rmse),
    rmse_smooth = map2_dbl(smooth_fits, test, rmse),
    rmse_wiggly = map2_dbl(wiggly_fits, test, rmse)
  )
```

Let's try to look at this better. 

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |>  # creates 2 columns (model and rmse)
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
# smooth model is best
```

## Child growth



```{r}
growth_df = 
  read_csv("data/nepalese_children.csv")
```

Weight v arm_c (arm circumference)

```{r}
growth_df |> 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5)
# has potential for non-linear association
```


Let's show the models we might use.

```{r}
growth_df =
  growth_df |> 
  mutate(
    # weight_cp is weight change point
    weight_cp7 = (weight > 7) * (weight - 7)
  )
```

Let's fit three models

```{r}
linear_mod = lm(armc ~ weight, data = growth_df)
pwl_mod = lm(armc ~ weight + weight_cp7, data = growth_df)
smooth_mod = mgcv::gam(armc ~ s(weight), data = growth_df)
```

```{r}
growth_df |> 
  add_predictions(linear_mod) |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5) + 
  geom_line(aes(y = pred), color = "red")
# working pretty well but need to fix something
```

```{r}
growth_df |> 
  add_predictions(pwl_mod) |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5) + 
  geom_line(aes(y = pred), color = "red")
# accounts for weight change at the beginning
```

```{r}
growth_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5) + 
  geom_line(aes(y = pred), color = "red")
# gives smoothed out end
```

Now cross validate!

```{r}
cv_df = 
  crossv_mc(growth_df, n = 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```


```{r}
cv_df = 
  cv_df |> 
  mutate(
    linear_mod = map(train, \(df) lm(armc ~ weight, data = df)),
    pwl_mod = map(train, \(df) lm(armc ~ weight + weight_cp7, data = df)),
    smooth_mod = map(train, \(df) mgcv::gam(armc ~ s(weight), data = df))
  ) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_pwl    = map2_dbl(pwl_mod, test, rmse),
    rmse_smooth = map2_dbl(smooth_mod, test, rmse)
  )
```

Create my boxplots!

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |> 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin()
# linear model underperforming compared to the other models
```



