---
title: "Linear Models"
output: github_document
---

```{r, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load key packages.

```{r}
library(tidyverse)
library(p8105.datasets)
```

Load the NYC airbnb data.

```{r}
data(nyc_airbnb)
```

Look at the data / do some cleaning.

```{r}
nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    stars = review_scores_location / 2
  ) |> 
  rename(
    borough = neighbourhood_group
  ) |> 
  filter(borough != "Staten Island") |> # not enough airbnbs in Staten Island to do regression
  select(price, stars, borough, room_type, neighbourhood)
```

Do regression!!!

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
fit # gives coefficient estimates
# R converts borough (a character variable) into a factors and turns it into alphabetical order
# reference group is Bronx here
```

do some additional cleaning then refit.

```{r}
nyc_airbnb |> 
  count(borough)

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(
    borough = fct_infreq(borough), # in order of how common (most appearances to least)
    room_type = fct_infreq(room_type)
  )

fit = lm(price ~ stars + borough, data = nyc_airbnb)
fit
# reference is Manhattan here
# intercept is the expected price of a rental in Manhattan when stars equal to 0
# Brooklyn is $50 cheaper than Manhattan
# Queens is $77 cheaper than Manhattan
```

Look at `lm` stuff

```{r, eval = FALSE}
summary(fit) # tells you about residuals, estimates, p-values, R^2
names(summary(fit))
summary(fit)[["coefficients"]] # pulls out coefficients table
summary(fit)[["df"]] # pulls out degrees of freedom

fitted.values(fit) # fitted values for all rows in the data frame
resid(fit)
```

Look at cleaner `lm` stuff

```{r}
fit |> 
  broom::tidy() |> 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) |> 
  select(term, estimate, p.value) |> 
  knitr::kable(digits = 3)
# nicely formatted table

fit |> 
  broom::glance() # gives important info you might need
```


## Diagnostics

Look at residuals.

```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> # computes residuals and adds it to the data frame
  modelr::add_predictions(fit) |> # gives predicted values
  filter(resid < 1000) |> 
  ggplot(aes(x = resid)) + 
  geom_histogram()
# histogram of residual values, most residuals centered around 0

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = borough, y = resid)) + 
  geom_violin()
# shows residual distribution across boroughs
# doesn't look like there's a constant variation across boroughs
# Manhattan has a wider distribution, more negative and highly skewed, others are increasingly narrowed

nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  filter(resid < 1000) |> 
  ggplot(aes(x = stars, y = resid)) + 
  geom_point()
# distribution of residuals according to stars
# less skewness and variability for airbnbs with less stars
```



## Hypothesis testing

```{r}
fit |> 
  broom::tidy()
# estimates for each individual coefficients
```

What about a categorical variable?

```{r}
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
# compare these two models (one has room type and other doesn't)

anova(fit_null, fit_alt) |> 
  broom::tidy()
# really small p-value tells us room type matters
```


## Interactions vs nested data


```{r}
# stars might be associated with borough, effect of stars differs from one borough to the next
# room type might also be associated with borough
# Interactions tell you how does one variable modify the main effect of another variable?
fit_interactions =
  lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)


fit_interactions |> 
  broom::tidy()
# affect of increasing stars in Brooklyn is 27.1 - 6.14
```


Let's just look at Brooklyn first.

```{r}
nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> # _ indicates data frame
  broom::tidy()
# affect of stars and room type in Brooklyn

nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Bronx") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

```


Write a short function

```{r}
# computes specific linear model based on what data is inputted
lm_airbnb = function(df) {
  
  lm(price ~ stars + room_type, data = df)
  
}

nyc_airbnb |> 
  filter(borough == "Queens") |> 
  lm_airbnb() |> # inputs data frame subset to Queens
  broom::tidy()
```

Create a list of dataframes, and iterate to fit the model each time.

```{r}
nested_lm_results = 
  nyc_airbnb |> 
  nest(data = -borough) |> # nest everything except borough, created column of data frames I care about
  mutate(
    fits = map(data, lm_airbnb), # fits column includes linear model fits for each borough
    results = map(fits, broom::tidy) # results column includes cleaned up data frames for lm fits for each borough
  ) |> 
  select(borough, results) |> 
  unnest(results)
```

Do some untidying.

```{r}
nested_lm_results |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term, 
    values_from = estimate
  )
# stars has biggest impact in Manhattan and Brooklyn
# affect of private room in Manhattan is more negative than the affect of private room in Bronx
# we don't have p-valuess so we can't tell if affect of stars is statistically significant across boroughs
```


Use an *anonymous* function instead of `lm_airbnb`

```{r}
# using a function that doesn't really exist, in the data frame
nested_lm_results = 
  nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)), # inputted the lm function here
    results = map(fits, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results)

nested_lm_results |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term, 
    values_from = estimate
  )

# use fct_reorder() if you want to recode the categorical variables such as borough
```



Let's do an even more extreme example!!

```{r}
manhattan_analysis =
  nyc_airbnb |> 
  filter(
    borough == "Manhattan"
  ) |> 
  nest(data = -neighbourhood) |> # separate dataset for each of the 32 neighborhoods
  mutate(
    fits = map(data, \(df) lm(price ~ stars + room_type, data = df)), # use anonymous function
    results = map(fits, broom::tidy)
  ) |> 
  select(neighbourhood, results) |> 
  unnest(results)
# fits linear model for each neighborhood in Manhattan (fits 32 models in total!)
```

Make a plot

```{r}
manhattan_analysis |> 
  filter(term == "stars") |> 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, estimate)
  ) |> 
  ggplot(aes(x = neighbourhood, y = estimate)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # rotate x-axis labels
# affect of increasing stars across neighborhoods in Manhattan
```


Example using binary outcomes...

```{r}
baltimore_df = 
  read_csv("data/homicide-data.csv") |> 
  filter(city == "Baltimore") |> 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) |> 
  select(resolved, victim_age, victim_race, victim_sex)
```

```{r}
fit_logistic = 
  baltimore_df |> 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = _, family = binomial()) 
```

```{r}
fit_logistic |> 
  broom::tidy() |> 
  mutate(OR = exp(estimate)) |>
  select(term, log_OR = estimate, OR, p.value) |> 
  knitr::kable(digits = 3)
```



